mlr_assg4

### 2022-09-21 hw4-1/abhinav/mlr.Rmd
Final Code push-1

new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
crime_rate_pred<-round(crime_rate_pred,2)
print(paste("Predicted Value:", crime_rate_pred))
```
# Interpreting the output of the linear regression model : 
Based on the model output , it appears that the features Ed, Ineq, Prob,M , Po1 are significant variables based on the p-values . 
* The lower the p-value , the more significant the variable  because the p-value signifies the probability of the estimated coefficients $\hat{\beta_{i}}'s$ coming from the underlying assumed distribution of null hypothesis (Null Hypothesis states that $\beta_{i}'s=0$). Thus , lower the p-value , higher the probability that the coefficient of the variable is not equal to 0 and hence that variable is a good predictor.
* For this model , the $R^{2}$ value is 0.8 which means that 80% of the variance in the dependent variable is explained by the set of these independent variables .
* The predicted crime_rate for the given data point is `r crime_rate_pred`.
* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place , error terms are normally distributed and residuals are homogeneous around the fitted line. However , we can see slight bends in the left and right end in the plot of residuals vs fitted values. Thus , a possible remedy can be thought of as modelling the log of dependent variable based on independent variable. (because taking a log transformation reduces the range of possible variables.)

@@

new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
crime_rate_pred<- exp(crime_rate_pred)
print(paste("Predicted Value:", round(crime_rate_pred,2)))
```
Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here . 
Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers.

>>>>

new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
crime_rate_pred<-round(crime_rate_pred,2)
print(paste("Predicted Value:", crime_rate_pred))
```
# Interpreting the output of the linear regression model : 
Based on the model output , it appears that the features Ed, Ineq, Prob,M , Po1 are significant variables based on the p-values . \
* The lower the p-value , the more significant the variable  because the p-value signifies the probability of the estimated coefficients $\hat{\beta_{i}}'s$ coming from the underlying assumed distribution of null hypothesis (Null Hypothesis states that $\beta_{i}'s=0$). Thus , lower the p-value , higher the probability that the coefficient of the variable is not equal to 0 and hence that variable is a good predictor.\

* For this model , the $R^{2}$ value is 0.8 which means that 80% of the variance in the dependent variable is explained by the set of these independent variables .\

* The predicted crime_rate for the given data point is `r crime_rate_pred`.\

* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place , error terms are normally distributed and residuals are homogeneous around the fitted line. However , we can see slight bends in the left and right end in the plot of residuals vs fitted values. Thus , a possible remedy can be thought of as modelling the log of dependent variable based on independent variable. (because taking a log transformation reduces the range of possible variables.)

@@

new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
crime_rate_pred<- exp(crime_rate_pred)
print(paste("Predicted Value:", round(crime_rate_pred,2)))
```
Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here .\ 
Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers.
